---
title: "RWorksheet_Hallares#4a"
author: "Marc Christian P. Hallares"
date: "2025-11-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
````

---

# 1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
shoe  <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 7.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5,
           13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 9.0, 9.0, 13.0)

height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0,
            71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0,
            72.0, 70.0, 69.0, 70.0)

gender <- c("F","F","F","F","M","F","F","M","M","F","M","F","M","M",
            "M","M","F","F","M","F","F","M","M","F","M","M","M","M")

df <- data.frame(ShoeSize = shoe, Height = height, Gender = gender)
df
```

---

## a. Describe the Data

The dataset has **28 observations** with the variables:

* **ShoeSize** – numeric
* **Height** – numeric (in inches)
* **Gender** – M or F

Shoe sizes range from **5.0 to 13.0**, and heights range from **59.0 to 77.0 inches**.

---

## b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.

```{r}
males   <- subset(df, Gender == "M")
females <- subset(df, Gender == "F")

males
females
```

---

## c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.

```{r}
mean_shoe   <- mean(df$ShoeSize)
mean_height <- mean(df$Height)

mean_shoe
mean_height
```

---

## d. Is there a relationship between shoe size and height? Why?

```{r}
cor(df$ShoeSize, df$Height)
```

A **positive correlation** indicates that taller individuals tend to have larger shoe sizes.

---

# 2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector

```{r}
months <- c("March","April","January","November","January","September","October",
            "September","November","August","January","November","November",
            "February","May","August","July","December","August","August",
            "September","November","February","April")
```

```{r}
factor_months_vector <- factor(months)
factor_months_vector
```

---

# 3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this case?

```{r}
summary(months)
summary(factor_months_vector)
```

---

# 4. Create a vector and factor for the table below.

```{r}
direction <- c("East", "West", "North")
frequency <- c(1, 4, 3)

factor_data <- data.frame(Direction = direction, Frequency = frequency)
factor_data
```

### Reordered Factor Levels

```{r}
new_order_data <- factor(factor_data$Direction,
                         levels = c("East", "West", "North"))
new_order_data
```

---

# 5. Enter the data below in Excel with file name = import_march.csv

## a. Import the excel file into the Environment Pane using read.table() function. Write the code.

```{r}
data_march <- read.table("import_march.csv",
                         header = TRUE,
                         sep = ",",
                         stringsAsFactors = FALSE)
```

## b. View the dataset. Write the R scripts and its result.

```{r}
data_march
# View(data_march)   # For RStudio
```

---

# 6. Full Search

```{r}
problem_6 <- function() {
  cat("Enter a number (1-50): ")
  number <- as.integer(readLines(con = "stdin", n = 1))

  if (number >= 1 && number <= 50) {
    cat("The chosen number is:", number, "\n")
  } else {
    cat("Out of Range\n")
  }
}

problem_6_test <- function(number) {
  if (number >= 1 && number <= 50) {
    cat("The chosen number is:", number, "\n")
  } else {
    cat("Out of Range\n")
  }
}

problem_6_test(20)
problem_6_test(55)
```

---

# 7. Change

```{r}
problem_7 <- function(price) {
  bills_1000 <- price %/% 1000
  remaining  <- price %% 1000

  bills_100 <- remaining %/% 100
  remaining <- remaining %% 100

  bills_50 <- remaining %/% 50

  total_bills <- bills_1000 + bills_100 + bills_50

  cat("Price of snack:", price, "pesos\n")
  cat("Minimum number of bills needed:", total_bills, "\n")
  cat("Breakdown:\n")
  cat("  1000 peso bills:", bills_1000, "\n")
  cat("  100 peso bills:", bills_100, "\n")
  cat("  50 peso bills:", bills_50, "\n")
}

problem_7(1250)
problem_7(350)
```

---

# The following is each student’s math score for one semester. Based on this, answer the following questions.

```{r}
names  <- c("Annie","Thea","Steve","Hanna")
grade1 <- c(85, 65, 75, 95)
grade2 <- c(65, 75, 55, 75)
grade3 <- c(85, 90, 80, 100)
grade4 <- c(100,90,85,90)

students <- data.frame(Name = names,
                       Grade1 = grade1,
                       Grade2 = grade2,
                       Grade3 = grade3,
                       Grade4 = grade4)

students
```

---

## a. Data Frame (Shown Above)

---

## b. Students with Average > 90 (No rowMeans)

```{r}
for (i in 1:nrow(students)) {
  avg <- (students$Grade1[i] + students$Grade2[i] +
          students$Grade3[i] + students$Grade4[i]) / 4

  if (avg > 90) {
    cat(students$Name[i], "'s average grade this semester is", avg, "\n")
  }
}
```

---

## c. Tests with Average < 80 (No mean)

```{r}
test_names <- c("Grade1","Grade2","Grade3","Grade4")

for (j in 2:5) {
  test_avg <- sum(students[, j]) / nrow(students)

  if (test_avg < 80) {
    cat("The", test_names[j-1], "test was difficult.\n")
  }
}
```

---

## d. Students Whose Highest Score > 90 (No max)

```{r}
for (i in 1:nrow(students)) {
  max_score <- students$Grade1[i]

  if (students$Grade2[i] > max_score) max_score <- students$Grade2[i]
  if (students$Grade3[i] > max_score) max_score <- students$Grade3[i]
  if (students$Grade4[i] > max_score) max_score <- students$Grade4[i]

  if (max_score > 90) {
    cat(students$Name[i], "'s highest grade this semester is", max_score, "\n")
  }
}
```
---
