---
title: "RWorksheet_Hallares#3a"
author: "Marc Christian P. Hallares"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1: Using Vectors

## a. First 11 letters
```{r vectors1}
LETTERS[1:11]
```

## b. Odd-numbered letters
```{r vectors2}
LETTERS[seq(1, 26, by = 2)]
```

## c. Vowels from LETTERS
```{r vectors3}
LETTERS[LETTERS %in% c("A", "E", "I", "O", "U")]
```

## d. Last 5 lowercase letters
```{r vectors4}
tail(letters, 5)
```

## e. Letters 15 to 24 (lowercase)
```{r vectors5}
letters[15:24]
```

---

# Exercise 2: Temperature Vectors

## a. Character vector of cities
```{r dataframe1}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
city
```

## b. Temperature vector
```{r dataframe2}
temp <- c(42, 39, 34, 34, 30, 27)
temp
```

## c. Create dataframe
```{r dataframe3}
weather <- data.frame(city, temp)
weather
```

## d. Rename columns
```{r dataframe4}
names(weather) <- c("City", "Temperature")
weather
```

## e. Print structure
```{r dataframe5}
str(weather)
```

## f. Show row 3 and 4
```{r dataframe6}
weather[3:4, ]
```

## g. Highest and lowest temperature
```{r dataframe7}
weather[which.max(weather$Temperature), ]
weather[which.min(weather$Temperature), ]
```

---

# Exercise 3: Matrices

## a. Create matrix
```{r matrix1}
m <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
m
```

## b. Multiply matrix by 2
```{r matrix2}
m2 <- m * 2
m2
```

## c. Row 2
```{r matrix3}
m2[2, ]
```

## d. Columns 3 and 4 of rows 1 and 2
```{r matrix4}
m2[1:2, 3:4]
```

## e. Columns 2 and 3 of row 3
```{r matrix5}
m2[3, 2:3]
```

## f. Column 4 only
```{r matrix6}
m2[, 4]
```

## g. Name rows and columns
```{r matrix7}
rownames(m2) <- c("isa", "dalawa", "tatlo")
colnames(m2) <- c("uno", "dos", "tres", "quatro")
m2
```

## h. Reshape original matrix
```{r matrix8}
m <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
dim(m) <- c(6, 2)
m
```

---

# Exercise 4: Arrays

## a. Create 3D array (values repeated twice)
```{r array1}
values <- rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), each = 2)
arr <- array(values, dim = c(2, 4, 3))
arr
```

## b. Number of dimensions
```{r array2}
dim(arr)
```

## c. Name rows, columns, and dimensions
```{r array3}
dimnames(arr) <- list(
  c("a", "b"),
  c("A", "B", "C", "D"),
  c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array")
)
arr
```
